04_DESeq_sexes

### 2019-08-27 analysis/04_DESeq2_sexes.Rmd
now this is actually allows comparison across sexes

```{r readfiles}
# import "colData" which contains sample information and "countData" which contains read counts
a.colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
a.countData <- read.csv("../results/00_counts.csv", header = T, row.names = 1)
geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)

# set levels
levels(a.colData$treatment)

a.colData$treatment <- factor(a.colData$treatment, levels = 
                             c("control", "bldg", "lay",
                               "inc.d3", "inc.d9", "inc.d17",
                               "hatch", "n5", "n9",

                               "m.inc.d3", "m.inc.d8", "m.inc.d9",
                               "m.inc.d17",  "m.n2"  ,
                                "prolong", "extend" ))

a.colData$sextissue <- as.factor(paste(a.colData$sex, a.colData$tissue, sep = "_"))

a.colData$lastday <- ifelse(grepl("m.inc.d3|m.inc.d9|m.inc.d17|m.n2", a.colData$treatment), "empty nest", 
                    ifelse(grepl("m.inc.d8|hatch|extend", a.colData$treatment),"chicks hatch",      
                     ifelse(grepl("n5", a.colData$treatment),"chicks early",
                     ifelse(grepl("n9", a.colData$treatment),"chicks later",
                     ifelse(grepl("control", a.colData$treatment),"control",
                     ifelse(grepl("bldg", a.colData$treatment),"nest building",
                     ifelse(grepl("prolong", a.colData$treatment),"eggs delay",
                      ifelse(grepl("lay", a.colData$treatment),"eggs lay",
                      ifelse(grepl("inc.d3", a.colData$treatment),"eggs early",
                     ifelse(grepl("inc.d9", a.colData$treatment),"eggs middle",
                    ifelse(grepl("inc.d17", a.colData$treatment),"eggs later", NA)))))))))))

a.colData$penultimate <-  ifelse(grepl("extend", a.colData$treatment),"eggs delay",  
                          ifelse(grepl("m.n2", a.colData$treatment),"chicks hatch",
                     ifelse(grepl("n5", a.colData$treatment),"chicks early",
                     ifelse(grepl("n9", a.colData$treatment),"chicks later",
                     ifelse(grepl("control", a.colData$treatment),"control",
                     ifelse(grepl("bldg|lay", a.colData$treatment),"nest building",
                     ifelse(grepl("prolong", a.colData$treatment),"eggs delay",
                      ifelse(grepl("inc.d3|m.inc.d3", a.colData$treatment),"eggs early",
                     ifelse(grepl("inc.d9|m.inc.d9|m.inc.d8", a.colData$treatment),"eggs middle",
                    ifelse(grepl("inc.d17|hatch", a.colData$treatment),"eggs later", NA))))))))))

a.colData$xlabel <- a.colData$treatment

levels(a.colData$xlabel) <-  c("control", "nest.building", "egg.lay",
                               "eggs.early", "eggs.early.remove", 
                               "eggs.mid", "eggs.mid.hatch", "eggs.mid.remove",
                               "eggs.end", "eggs.end.remove",
                               "chicks.hatch",  "chicks.hatch.remove"  ,
                               "chicks.mid", "eggs.delay", "eggs.delay.hatch", "chicks.end")

a.colData$lastday <- factor(a.colData$lastday, levels =  c("control", "nest building", "eggs lay",
                               "eggs early", "eggs middle",  "eggs later","eggs delay",
                               "chicks hatch", "chicks early",  "chicks later", "empty nest"))

a.colData$penultimate <- factor(a.colData$penultimate, levels =  c("control", "nest building",
                               "eggs early",  "eggs middle",  "eggs later",  "eggs delay",
                               "chicks hatch", "chicks early",  "chicks later"))

summary(a.colData[c(7,3,4,5,8,9, 10,11)])
```

@@

```{r dds, cache = T}
dds.female_hypothalamus <- subsetDESeq(a.colData, a.countData, "female_hypothalamus")
dds.female_pituitary <- subsetDESeq(a.colData, a.countData, "female_pituitary" )
dds.female_gonad <- subsetDESeq(a.colData, a.countData, "female_gonad" )
dds.male_hypothalamus <- subsetDESeq(a.colData, a.countData, "male_hypothalamus" )
dds.male_pituitary <- subsetDESeq(a.colData, a.countData, "male_pituitary"  )
dds.male_gondad <- subsetDESeq(a.colData, a.countData, "male_gonad")
```

>>>>

```{r readfiles}
# import "colData" which contains sample information and "countData" which contains read counts
c.colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
c.countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)

# set levels
c.colData$treatment <- factor(c.colData$treatment, levels = 
                              c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9"))
levels(c.colData$treatment)
c.colData$sextissue <- as.factor(paste(c.colData$sex, c.colData$tissue, sep = "_"))
summary(c.colData[c(7,3,4,5,8)])
```

@@

```{r dds}
dds.hypothalamus <- subsetDESeq2(c.colData,  c.countData, c("female_hypothalamus","male_hypothalamus") )
dds.pituitary <- subsetDESeq2(c.colData,  c.countData, c("female_pituitary", "male_pituitary" ) )
dds.gonad <- subsetDESeq2(c.colData,  c.countData, c("female_gonad", "male_gonad") )
```

```{r pca}
hypPCA <- returnPCAs2(dds.hypothalamus)
plotPC12(hypPCA, "female and male hypothalamus")

pitPCA <- returnPCAs2(dds.pituitary)
plotPC12(pitPCA, "female and male pituitary")

gonPCA <- returnPCAs2(dds.gonad)
plotPC12(gonPCA, "female and male gonad")
```
